---
title: "Draft Data Visualization"
author: "Lê Khánh Nhân"
date: "20-2-2026"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# 1. Load Libraries
library(tidyverse)
library(reshape2)
dataset_3D_printer <- read_csv("data.csv")
temp_data <- dataset_3D_printer %>%
  mutate(
    infill_pattern = as.numeric(infill_pattern != "grid"),
    material = as.numeric(material != "abs")
  )
```

## Overview information

Overview of 3D-Printing parameters and outputs within the data set:

```{r summary, echo=TRUE}
glimpse(dataset_3D_printer)
summary(dataset_3D_printer)
```
## Histogram

Histogram of each parameters:
```{r}
temp_data <- dataset_3D_printer %>%
  mutate(infill_pattern = as.numeric(infill_pattern != "grid"),
         material = as.numeric(material != "abs"))

gathered_data <- gather(temp_data)
gathered_data <- gathered_data %>%
  mutate(key = factor(key, levels = unique(gathered_data$key)))

ggplot(data = gathered_data, aes(x = value)) +
  geom_histogram() +
  facet_wrap(~ key, scales = "free", nrow = 4) +
  labs(x = NULL, y = NULL)
```



## Pearson correlation matrix

By creating a Pearson correlation matrix, we can observe the relationships between 
different parameters and between inputs and outputs.
By that, we can choose what set of parameters we should look deeper into.

```{r Pearson Correlation Matrix}

# Correlation Matrix Logic
cormat <- round(cor(temp_data), 2)

# Helper function to get upper triangle (removes the mirror image)
get_upper_tri <- function(cormat) {
  cormat[lower.tri(cormat)] <- NA
  return(cormat)
}

# Helper function to reorder by similarity (clustering)
reorder_cormat <- function(cormat) {
  dd <- as.dist((1 - cormat) / 2)
  hc <- hclust(dd)
  cormat[hc$order, hc$order]
}

# Apply reordering and extract upper triangle
cormat <- reorder_cormat(cormat)
upper_tri <- get_upper_tri(cormat)

# Visualize with ggplot2
melted_cormat <- melt(upper_tri, na.rm = TRUE)

ggplot(melted_cormat, aes(Var2, Var1, fill = value)) +
  geom_tile(color = "black") +
  scale_fill_gradient2(
    low = "blue", high = "red", mid = "azure", 
    midpoint = 0, limit = c(-1, 1), 
    name = "Pearson\nCorrelation"
  ) +
  geom_text(aes(label = value), color = "black", size = 3.5) +
  coord_fixed() +
  theme_minimal() +
  labs(title = "3D Printer Parameters Correlation Matrix", x = NULL, y = NULL) +
  theme(
    axis.text.x = element_text(angle = 45, vjust = 1, size = 10, hjust = 1),
    panel.grid.major = element_blank(),
    panel.border = element_blank(),
    panel.background = element_blank(),
    axis.ticks = element_blank(),
    legend.justification = c(1, 0),
    legend.position = c(0.6, 0.7),
    legend.direction = "horizontal"
  ) +
  guides(fill = guide_colorbar(
    barwidth = 7, barheight = 1,
    title.position = "top", title.hjust = 0.5
  ))



```
We can observe that 
1. Highly correlated parameters like Layer height to the roughness level, nozzle temperature to tension strenth, infill density to tension strength, material to elongation, we will analyze those deeper later on.
2. Temperature related parameters highly correlated to each other as espected like bed_temperature, nozzle temperature, fanspeed.
3. Infill pattern hardly affect the final product
4. Printing material is more important than infill pattern
...

## Layer height to Roughness

Layer height and roughness have a positive correlation with both materials:

```{r Layer Height to Roughness, echo=TRUE}
ggplot(dataset_3D_printer, aes(x = layer_height, 
                               y = roughness,
                               color = material)) +
geom_point(shape = 18) + 
geom_smooth(method = "lm", se = FALSE) + 
scale_color_brewer(palette = "Set1") + 
labs(x = "Layer height", 
       y = "Roughness",
       color = "Material",
       title = "Correlation between Layer Height and Roughness",
       caption = "2011-2025 The Bootstrap Authors")

```

We can conclude that greater layer height will make the surface more rough.

## Tension Strength to Nozzle temperature

Tension strength with different nozzle temperature
```{r Tension Strength and Nozzle temperature}
ggplot(dataset_3D_printer, aes(x = nozzle_temperature, 
                               y = tension_strenght,
                               color = material)) +
  geom_point(shape = 18) + 
  geom_smooth(method = "lm", se = FALSE) + 
  scale_color_brewer(palette = "Set1") + 
  labs(x = "Nozzle temperature", 
       y = "Tension strength",
       color = "Material",
       title = "Correlation between Nozzle temperature and Tension strength",
       caption = "2011-2025 The Bootstrap Authors")


```
Pla tend to perform better when printed at a higher nozzle temperature 
Abs is a complete oposite and will perform much worse when printed at a higher nozzle temperature.

## Infill density to Tension strength
```{r}
ggplot(dataset_3D_printer, aes(x = infill_density, 
                               y = tension_strenght,
                               color = material)) +
  geom_point(shape = 18) + 
  geom_smooth(method = "lm", se = FALSE) + 
  scale_color_brewer(palette = "Set1") +
  labs(x = "Infill density", 
       y = "Tension strength",
       color = "Material",
       title = "Correlation between Infill density and Tension strength",
       caption = "2011-2025 The Bootstrap Authors")
```
Printed part with more infill density can perform better when applied tensile force, especially pla.


## Material to Elongation
```{r}

ggplot(dataset_3D_printer, aes(x = material,
                               y = elongation,
                               color = infill_pattern)) +
  geom_boxplot(shape = 19) + 
  geom_smooth(method = "lm", se = FALSE) + 
  scale_color_brewer(palette = "Set2") +
  labs(x = "Material", 
       y = "Elongation",
       color = "Infill pattern",
       title = "Correlation between material and elongation",
       caption = "2011-2025 The Bootstrap Authors")


```

Pla can withstand more elongation compare to abs.